---
title: "compet"
author: "Shriya Kapavarapu"
date: "2024-07-14"
output: html_document
---

```{r}
b = 3
d = 0.5
dd = (3*0.7 -1) / (2 * 0.7 * K)
t = 1:n

exp.simu3 = rep(NaN,n)

#assigns first value in vector to 100
exp.simu3[1] <- 100
exp.simu3[1]

```

```{r}
for( i in 2:length(exp.simu3)){
  births = exp.simu3[i-1] * b
  recruitment = (births * (d / (1+ (dd * births))))
  deaths = exp.simu3[i-1] * d
  exp.simu3[i] = exp.simu3[i-1] + recruitment - deaths
}
```

```{r}
sim_df$pop3 = exp.simu3

```
#1. You will take the last population model created during the live coding demo and choose one element to become stochastic. What this means, is instead of staying constant for the entire simulation, you will use a probability distribution to add variation in each time step. You may choose one of the following parameters:

#Number of births

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(gam)

```

```{r}
b = 3
d = 0.5
dd = (3*0.7 -1) / (2 * 0.7 * K)

exp.simu4 = rep(NaN,n)

#assigns first value in vector to 100
exp.simu4[1] <- 100
exp.simu4[1]

```

```{r}
for( i in 2:length(exp.simu4)){
  births = exp.simu4[i-1] * rpois(n=1,b)
  recruitment = (births * (d / (1+ (dd * births))))
  deaths = exp.simu4[i-1] * d
  exp.simu4[i] = exp.simu4[i-1] + recruitment - deaths
  
}
```



```{r}
sim_df$pop4 = exp.simu4

ggplot(sim_df)+
  geom_point(aes(t,pop4), color = "purple")+
  geom_line(aes(t,pop4), linetype = "dashed")+
  theme_bw()+
  labs(x = "Time", y = "Population size")
```
```{r}

```
#3. Run a paired t-test to see if there are significant differences between the model without stochasticity (from the demo) and with stochasticity. Check the assumptions. What test could you perform instead if the assumptions are not met? (10 points)
```{r}
t.test(exp.simu3, exp.simu4, paired = TRUE, data = "two.sided")
``` 

Our p-test was smaller than the standard 0.05 value, meaning that our data is significant. The null hypothesis is rejected, since the mean difference of the two graphs did not equal zero. 
# -------------------------------------------------------------------------------------------------
Part 2 

You will take what you learned from the simulation of population growth and create a simulation for either competition or predation

The equations for these are provided in your lecture 4 slides. You can choose your own parameter values (for things such as r, competitive ability, etc. but they must be realistic). The key difference is you will need to keep track of population values for 2 populations instead of 1. (20 points)

Competition: 

```{r}
exp.simu5 = rep(NaN, n)

#assigns the first value in vector to 100
exp.simu5[1] <- 100
exp.simu5[1]

r1 = 0.78
K1 = 786

exp.simu6 = rep(NaN, n)
#assigns the first value in vector to 100
exp.simu6[1] <- 100
exp.simu6[1]

r2 = 0.45
K2 = 1338
```

```{r}
delta_t <- 1
alpha1 <- -0.5 
alpha2 <- -0.3

sim_df <- data.frame( t = 1:50)

for(i in 2:length(exp.simu5)) {
  delta_N1 <- (exp.simu5[i-1] *r1) *((K1-exp.simu5[i-1]+(exp.simu6[i-1]*alpha1))/ K1) * delta_t
  exp.simu5[i] <- exp.simu5[i-1] + delta_N1
  
  delta_N2 <- (exp.simu6[i-1] *r2) *((K2-exp.simu6[i-1]+(exp.simu5[i-1]*alpha2))/ K2) * delta_t
  exp.simu6[i] <- exp.simu6[i-1] + delta_N2
}

sim_df$pop5 = exp.simu5
sim_df$pop6 = exp.simu6

ggplot(sim_df) +
  geom_point(aes(t , pop5), color = "red") +
  geom_line(aes(t , pop5), linetype = "dashed") +
  geom_point(aes(t , pop6), color = "blue") +
  geom_line(aes(t , pop6), linetype = "dashed") +
  theme_bw()+
  labs(x = "Time" , y = "Population Size")

```
```{r}
exp.simu7 = rep(NaN, n)

#assigns the first value in vector to 100
exp.simu7[1] <- 100
exp.simu7[1]

r3 = 0.78
K3 = 786

exp.simu8 = rep(NaN, n)
#assigns the first value in vector to 100
exp.simu8[1] <- 100
exp.simu8[1]

r4 = 0.45
K4 = 1338
```

```{r}
delta_t <- 1
alpha3 <- -0.5
alpha4 <- -0.9

sim_df <- data.frame( t = 1:50)

for(i in 2:length(exp.simu7)) {
  delta_N3 <- (exp.simu7[i-1] *r1) *((K1-exp.simu7[i-1]+(exp.simu8[i-1]*alpha3))/ K1) * delta_t
  exp.simu7[i] <- exp.simu7[i-1] + delta_N3
  
  delta_N4 <- (exp.simu8[i-1] *r2) *((K2-exp.simu8[i-1]+(exp.simu7[i-1]*alpha4))/ K2) * delta_t
  exp.simu8[i] <- exp.simu8[i-1] + delta_N4
}

sim_df$pop7 = exp.simu7
sim_df$pop8 = exp.simu8

ggplot(sim_df) +
  geom_point(aes(t , pop7), color = "green") +
  geom_line(aes(t , pop7), linetype = "dashed") +
  geom_point(aes(t , pop8), color = "purple") +
  geom_line(aes(t , pop8), linetype = "dashed") +
  theme_bw()+
  labs(x = "Time" , y = "Population Size")

```
#Write a short reflection (~100 words) explaining the results of changing parameter values on your two populations. (10 points)

In the second simulation of competition, I changes the alpha values of the two 